---
title: "Find Genes Documentation/ Tutorial"
author: "Hallie Arno"
format: 
  html:
    embed-resources: true
editor: source
---

This function replaces bedtools intersect for finding which genes SNPs are in for Atlantic salmon and Brown trout. It uses .csv files of the position (start and end) of each gene. It then merges the input SNP dataframe with the list of all genes, to return which genes contain the SNPs of interest. 

Below is an example ofhow to use the function. 

First need to source in file, make sure the entire folder is kept. One issue may be that the entire folder is not in the working directory. This can be fixed easily by making a copy of the entire folder (geneMatching) and putting it as a subdirectory in your working directory. The function requires three packages to be installed: dplyr, data.table, and this.path.

```{r}
#| message: false
#| warning: false

library(dplyr) 

source("~/Desktop/Projects/geneMatching/findGenes.R")
```


To test the function, I made a dataframe with some made-up SNP positions. This is a similar format to how PLINK exports results (such as FST or GWAS), but column names may be different. 
```{r}
chroms <- c(1, 5, 16, 24, 26)
position <- c(10000, 750000, 1300, 40000, 88600)
mySNPs <- data.frame(cbind(chroms, position)) %>% 
  mutate(ID = paste(chroms, position, sep = "_"))

rmarkdown::paged_table(mySNPs)
```

Now to use the function to see if any of the made up SNPs are in genes. 
Note that the column names and species ("salmon" or "trout") must be specified:
```{r}
myGenes <- findGenes(snpsDf = mySNPs, pos = "position", chrom = "chroms", ID = "ID", species = "salmon")
rmarkdown::paged_table(myGenes)
```
Now these can be merged back with results (such as FST or GWAS), plotted as a manhattan plot, or used for a Gene Ontology Enrichment analysis. 



